Resilience and Home-Space for WSTS \thanks{This work was partly done while the authors were supported by the Agence Nationale de la Recherche grant BraVAS (ANR-17-CE40-0028).}

### 2023-11-02 VMCAI-revised-version/final.tex
up

\problemx{zero-reachability}
{A reset-VASS $V=(Q,T)$ of dimension $d$, and $p(\textbf{u}) \in Q \times \N^d$.} 
{  $\exists q \in Q, ~ p(\textbf{u}) \to^* q(\textbf{0})$ ?\\}

% 
{\sc Resilience} (and {\sc $t$-liveness}) needs a set (resp. a transition) to always remains reachable (resp. fireable), while
{\em zero-reachability} concerns itself only with the existence of at least one way to reach some state $q(\textbf{0})$,
which could be unreachable from other reachable states. 
%
Despite this it is possible to reduce 
the {\em zero-reachability problem} 
to
{\sc $t$-liveness}.
% The following Proposition is a variation  to reset-VASS of Theorem~5.5 in \cite{peterson1981petri} originally stated for Petri nets,
The following Proposition is a variation  to reset-VASS 
of~\cite[Theorem~5.5]{peterson1981petri} originally stated for Petri nets,
 whose proof can be seen in Appendix~\ref{appendix}%
.
\begin{proposition}\label{liveness reset}
The  zero reachability problem can be reduced to {\sc $t$-liveness}.
\end{proposition}

@@

We now consider the case $\Safe = \mathop{\downarrow} \Safe$.
%
Unfortunately {\sc State-resilience} is undecidable in this case. 
%\mathieu{GOTO}
Unfortunately {\sc State-resilience} is undecidable in this case. As for {\sc Resilience} in WSTS when $\Safe = \downarrow \Safe$, this stems from undeciability of the corresponding problem in Minsky machines, the executions of which can be simulated by reset-VASS, as seen in 
Section~\ref{safe-down} and developped more in Appendix~\ref{minsk}.

@@

@@

\problemx{zero-reachability}
{A $d$-reset-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$.} 
{  $\exists q \in Q, ~ p(\textbf{u}) \to^* q(\textbf{0})$ ?\\}

@@

\noindent
\textbf{Proposition~\ref{liveness reset}.}
{\em In reset Petri nets, the zero reachability problem can be reduced to {\sc $t$-liveness}.}\\

@@

\noindent
\textbf{Corollary}
{\em In reset-VASS, the zero reachability problem can be reduced to {\sc $t$-liveness}.}

@@

\begin{proof}
First, note that deciding whether from all inputs, all runs are terminating
 is undecidable for Minsky machines with at least $2$ counters, and, in particular, for deterministic ones as well~\cite{Min61}. Hence 
  deciding whether from all inputs, there exists at least one terminating run
 is undecidable for Minsky machines with at least $2$ counters, since,
 for deterministic Minsky machines, the two problems are one and the same.

@@

\iffalse
\section{Structural halting in reset-VASS}

We claim {\sc Structual existential halting} in reset-VASS is undecidable.  
This is based on the construction providng the undecidability of structural termination in Lossy counter machines and in reset-VASS~\cite{dufourd1998reset,mayr1998lossy}. Strictly speaking undecidability of
{\sc Structual existential halting} is not considered in
\cite{mayr1998lossy},
but 
the encoding of Minsky machines can be reused
since 
 {\sc Structual existential halting}
 is equivalent to
 structural termination in the case of deterministic Minsky machines and is thus
 undecidable for Minsky machines in general.
\fi

>>>>

\problemx{zero-reachability}
{A reset-VASS $V=(Q,T)$ of dimension $d$, $q_f \in Q$, $p(\textbf{u}) \in Q \times \N^d$.} 
{  $%\exists q \in Q, ~ 
p(\textbf{u}) \to^* q_f(\textbf{0})$ ?\\}

% 
{\sc Resilience} (and {\sc $t$-liveness}) needs a set (resp. a transition) to always remains reachable (resp. fireable), while
{\em zero-reachability} concerns itself only with the existence of at least one way to reach some state $q_f(\textbf{0})$,
which could be unreachable from other reachable states. 
%
Despite this it is possible to reduce 
the {\em zero-reachability problem} 
to
{\sc $t$-liveness}.
% The following Proposition is a variation  to reset-VASS of Theorem~5.5 in \cite{peterson1981petri} originally stated for Petri nets,
The following Proposition is a variation  to reset-VASS 
of~\cite[Theorem~5.5]{peterson1981petri} originally stated for Petri nets,
 whose proof can be seen in Appendix~\ref{appendix}%
.
\begin{proposition}\label{liveness reset}
The  zero-reachability problem can be reduced to {\sc $t$-liveness}.
\end{proposition}

@@

We now consider the case $\Safe = \mathop{\downarrow} \Safe$.
%
Unfortunately {\sc State-resilience} is undecidable in this case. 
%\mathieu{GOTO}
This stems from undeciability of zero-reachability in reset-VASS.

@@

\begin{proof}
Let $V$ be a reset-VASS and $q_0({\bf 0})$ an initial state of $V$,
we build $V'$ from $V$ by additionnally allowing from any control-state to get back to 
$q_0({\bf 0})$
by a transition reseting every counter and
changing the control-state to $q_0$.
Then a state $q_f({\bf 0})$ is reachable from $q_0({\bf 0})$ in $N$
iff $q_f({\bf 0})$ is reachable from every element of $\post^*(q_0({\bf 0}))$ in $N'$,
which corresponds to {\sc State-resilience}  for $\Safe =  \{ q_f({\bf 0}) \}$
and $s_0 = q_0({\bf 0})$.
\end{proof}

@@

\problemx{zero-reachability}
{A $d$-reset-VASS $V=(Q,T)$, $q_f \in Q$ and $p(\textbf{u}) \in Q \times \N^d$.} 
{  $%\exists q \in Q, ~ 
p(\textbf{u}) \to^* q_f(\textbf{0})$ ?\\}

@@

\noindent
\textbf{Proposition~\ref{liveness reset}.}
{\em In reset Petri nets, the zero-reachability problem can be reduced to {\sc $t$-liveness}.}\\

@@

\noindent
\textbf{Corollary}
{\em In reset-VASS, the zero-reachability problem can be reduced to {\sc $t$-liveness}.}

@@

\begin{proof}
First, note that deciding whether from all inputs, all runs are terminating
 is undecidable for Minsky machines with at least $2$ counters, and, in particular, for deterministic ones as well. Hence 
  deciding whether from all inputs, there exists at least one terminating run
 is undecidable for Minsky machines with at least $2$ counters, since,
 for deterministic Minsky machines, the two problems are one and the same.

@@
